---
title: "PM566 Final"
author: "Changqing Su"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE

---

```{r setup, message=FALSE, echo=FALSE, warning=FALSE}

library(data.table)
library(tidyverse)
library(dplyr)
library(plotly)
library(DT)
library(knitr)

# Initialize code chunk options
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  eval=TRUE,
  echo = F,
  cache = FALSE,
  fig.width = 7, 
  fig.align = 'center',
  fig.asp = 0.618,
  out.width = "700px",
  class.source = "code-r")
```

<br>

# Introduction

At the end of 2019, a novel coronavirus was identified as the cause of a cluster of pneumonia cases in China. It rapidly spread, and resulted in an epidemic throughout the world. In February 2020, the World Health Organization designated the disease COVID-19, which stands for coronavirus disease 2019. This report is designed to explore the association between COVID-19 death and state, sex, age groups.

<br>

# Methods

<br>

The data was obtained from the US CDC, Centers of Disease Control and Prevention on Oct 17, 2020. The data include deaths due to coronavirus disease 2019 (COVID-19), pneumonia, and influenza reported to NCHS by sex, age group and state. \
There age group variable contains overlapping data. Therefore, only portion of the data was used. \
The proportionate mortality ratio (PMR) was obtained by using COVID-19 deaths/total deaths. \

<br>

# Preliminary Results

<br>

```{r, warning=FALSE,message=FALSE,echo = FALSE}
source("Mid_data.R")
```

```{r midtermplots, class.source="code-r-small"}

figure1=covid_2  %>%
plot_ly(x = ~`Age group`, y = ~PMR, 
        type = 'scatter', mode = 'markers', color = ~State,
        size = ~`Total Deaths`, sizes = c(2, 30), 
         marker = list(sizemode='diameter', opacity=0.6)) %>%
  layout(title = "Age groups vs PMR by states",
                  yaxis = list(title = "Proportionate Mortality Ratio"), xaxis = list(title = "Age groups"))
  

figure2= covid_2 %>%
  plot_ly(x=~State , y = ~`COVID-19 Deaths`, color = ~Sex) %>%
  layout(title = "State vs COVID-19 Deaths by Sex",
                  yaxis = list(title = "COVID-19 Deaths"), xaxis = list(title = "State"))
  

figure3 =covid_2 %>%
  plot_ly(x=~`Age group` , y = ~PMR, color = ~Sex) %>%
  add_bars() %>%
  layout(title = "Age group vs PMR by Sex",
                  yaxis = list(title = "Proportionate Mortality Ratio"), xaxis = list(title = "Age group "),barmode = "stack") 




g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  lakecolor = toRGB('white')
)

figure4=plot_geo(covid_state) %>%
  add_trace(
    z = ~PMR, text = ~State, span = I(0),
    locations = state.abb, locationmode = 'USA-states' , color="Reds") %>% 
  layout(title='Covid PMR Map',
         geo = g)


```

### Covid-19 in State Level {.tabset}

#### Figure 1

<br>

```{r PMR within different age groups by states, echo=FALSE}
figure4
```

#### Table 1

```{r , echo=FALSE}
knitr::kable(head(covid_simp))
```

### {-}

According to Figure 1, New York has the highest proportionate mortality ratio of COVID-19. This may be caused by small living space per capita. Since the virus can spread from an infected person’s mouth or nose in small liquid particles, others are very easily get infected when staying with infected people. \
According to Table 1, California also has a large number of deaths due to Covid-19, but California has a very low PMR comparing to New York. This may be because that the state government did a better job on stopping the spread of Covid-19.


<br>

### Figure 2

<br>

```{r Covid deaths within different sex by states, echo=FALSE}
figure2
```

<br>

According to Figure 2, New York, California, Florida, and Texas shows that more males died from Covid-19 than females. Other states had close numbers between males and females. This may be because that males are more likely to have social communications with others, and they have a larger chance to contact with infect people.

<br>

### Figure 3

<br>

```{r PMR within different sex by age groups, echo=FALSE}
figure3
```

<br>

According to Figure 3, the Proportionate Mortality Ratio increases as the age increases. Older people are more likely to die due to Covid-19. Sex does not affect Proportionate Mortality Ratio of Covid-19 in this plot.

<br>

# Conclusion

In conclusion, the number of deaths due to Covid-19 varies in different states, sex, and age groups. \
The New York State has the highest proportionate mortality ratio due to COVID-19. Generally more people die from COVID-19 in New York State. \
Slightly more males died from COVID-19 than females. But the PMR does not vary between males and females. \
In addition, older people (greater than 65 years old) are more likely to die (more than 8%) because of COVID-19. 
